Supplementary Information

### 2016-12-02 SupplementaryInformation.Rmd
Finished read-through of main text, pushing edits

# Climate changes between 1895-1924 and 1950-1980  (vs. main paper analysis of 1895-1924 and 1990-2014)

If a 1950-1980 climate normal is used instead of 1990-2014 climate normal, then between 1895-1924 and 1950-1980, the regional climate changes are as follows: ($T_{min}$) increased 0.5^o^C between 1895-1924 and 1950-1980, ($T_{max}$) increased by 0.35^o^C, while T_diff changed by -0.15^o^C

@@

<!-- COM: ST4: scientific names needed -->
<!-- COM: ST4: Correct column names and formatting for climate variables.  e.g. ppt should be PAnn -->
<!-- COM: Correct formatting for variable names in text.  I have tried to do this where I have seen them, but may have missed some variables.   -->
**Supplemental Table 4**: As Table 1 in the main text, except that the compounding (□), confounding (-), and counteracting (○) assessments are based on a modern climate normal for 1950-1980 instead of 1990-2014. The results for ($T_{min}$), ($T_{diff}$), and ($P_{ann}$) are unchanged. For ($T_{max}$), the sign of the climate change flips from negative (cooling) to positive (warming), indicating that the apparent directionality of ($T_{max}$) change is sensitive to choice of modern climate normal.  (Note that directionality of ($T_{min}$) is unchanged, consistent with 20^th^- and 21^st^-century warming).   See also Supplemental Figures 3, 4.

@@

**Supplemental Figure 3**:  As Figure 4 in the main text, except that the modern climate normal is for 1950-1980 instead of 1990-2014. The results for ($T_{min}$) and ($P_{ann}$) are largely unchanged between this analysis and that shown in Fig. 4, indicating that the finding of tree-climate distribution shifts is robust to choice of modern cliamte normal.  However, ($T_{max}$) shows much closer overlap between the PLS and FIA periods, suggesting that the reported difference in tree-climate distributions for ($T_{max}$) is sensitive to choice of modern climate normal.

@@

**Supplemental Figure 4**:  As Figure 5 in the main text, except that the modern climate normal is for 1950-1980 instead of 1990-2014.  Few differences are apparent between this figure and Figure 5, suggesting that the attribution of tree-climate distribution shifts to changes in climate versus changes in vegetation is insensitive to choice of modern climate normal.

>>>>

# Climate changes between 1895-1924 and 1950-1979  (vs. main paper analysis of 1895-1924 and 1990-2014)

If a 1950-1979 climate normal is used instead of 1990-2014 climate normal, then between 1895-1924 and 1950-1979, the regional climate changes are as follows: ($T_{min}$) increased 0.5^o^C between 1895-1924 and 1950-1979, ($T_{max}$) increased by 0.35^o^C, while T_diff changed by -0.15^o^C

@@

<!-- COM: ST4: scientific names needed -->
<!-- COM: ST4: Correct column names and formatting for climate variables.  e.g. ppt should be PAnn -->
<!-- COM: Correct formatting for variable names in text.  I have tried to do this where I have seen them, but may have missed some variables.   -->
<!-- COM: For TMax, why does the sign of the vegetation change flip from negative to positive between Table 1 and this Supp Table 4?  I can understand why the sign of the climate change would alter as we play around with different climate normals, but the sign of the vegetation change should be stable.  The other climate variables seem to be unaffected.  This may be the bug you mentioned by email.   -->
**Supplemental Table 4**: As Table 1 in the main text, except that the compounding (□), confounding (-), and counteracting (○) assessments are based on a modern climate normal for 1950-1979 instead of 1990-2014. The results for ($T_{min}$), ($T_{diff}$), and ($P_{ann}$) are unchanged. For ($T_{max}$), the sign of the climate change flips from negative (cooling) to positive (warming), indicating that the apparent directionality of ($T_{max}$) change is sensitive to choice of modern climate normal.  (Note that directionality of ($T_{min}$) is unchanged, consistent with 20^th^- and 21^st^-century warming).   See also Supplemental Figures 3, 4.

@@

**Supplemental Figure 3**:  As Figure 4 in the main text, except that the modern climate normal is for 1950-1979 instead of 1990-2014. The results for ($T_{min}$) and ($P_{ann}$) are largely unchanged between this analysis and that shown in Fig. 4, indicating that the finding of tree-climate distribution shifts is robust to choice of modern cliamte normal.  However, ($T_{max}$) shows much closer overlap between the PLS and FIA periods, suggesting that the reported difference in tree-climate distributions for ($T_{max}$) is sensitive to choice of modern climate normal.

@@

**Supplemental Figure 4**:  As Figure 5 in the main text, except that the modern climate normal is for 1950-1979 instead of 1990-2014.  Few differences are apparent between this figure and Figure 5, suggesting that the attribution of tree-climate distribution shifts to changes in climate versus changes in vegetation is insensitive to choice of modern climate normal.
